No time for dead time - Use the Fourier amplitude differences to normalize dead time-affected periodograms

### 2017-12-10 paper/deadtime.tex
Complete draft of the revised paper - MB

\newcommand{\given}{\ensuremath{\,|\,}}
\newcommand{\dd}{\ensuremath{\mathrm{d}}}
\newcommand{\counts}{\ensuremath{y}}
\newcommand{\pars}{\ensuremath{\theta}}
\newcommand{\mean}{\ensuremath{\lambda}}
\newcommand{\likelihood}{\ensuremath{{\mathcal L}}}
\newcommand{\Poisson}{\ensuremath{{\mathcal P}}}
\newcommand{\Uniform}{\ensuremath{{\mathcal U}}}

\newcommand{\Normal}{\ensuremath{{\mathcal N}}}
\newcommand{\bg}{\ensuremath{\mathrm{bg}}}
\newcommand{\word}{\ensuremath{\phi}}

@@

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
Dead time affects many of the instruments used in X-ray astronomy, by producing a strong distortion in power density spectra. This can make it difficult to model the aperiodic variability of the source or look for quasi-periodic oscillations.
Whereas in some instruments a simple a-priori correction for dead time-affected power spectra is possible, 
this is not the case for others such as \nustar, where the dead time is non-constant and long ($\sim$2.5\,ms).
\citet{Bachetti+15} suggested the cospectrum obtained from light curves of independent detectors within the same instrument as a possible way out, but this solution has always only been a partial one: the measured rms was still affected by dead time, because the width of the power distribution of the cospectrum was modulated by dead time in a frequency-dependent way.

@@

When using data from missions carrying two or more \textit{identical and independent} detectors like \nustar, \citet{Bachetti+15} proposed an approach to mitigate instrumental effects like dead time exploiting this redundancy: 
where in standard analysis, light curves of multiple detectors are summed before Fourier transforming the summed light curve, 
it is possible to instead Fourier-transform the signal of two independent detectors and combine the Fourier amplitudes in a \textit{cospectrum} -- the real part of the cross spectrum -- instead of the periodogram. 
Since dead time is uncorrelated between the two detectors, the resulting powers have a mean white noise level fixed to 0, which resolves the first and most problematic issue created by dead time (see details in \citealt{Bachetti+15}); however, the resulting powers no longer follow the statistical distribution expected for power spectra, and their probability distribution is frequency-dependent.
Whereas a noise cospectrum in the absence of dead time would follow a Laplace distribution (Huppenkothen and Bachetti, sub.),
dead time affects the width of the probability distribution for cospectral powers and modulates the measured rms proportionally to the distortion acted on power spectra.
In this Letter, we show a method to precisely recover the shape of the power density spectrum by looking at the difference of the Fourier amplitudes of the light curves of two independent detectors.
This difference, in fact, contains information on the uncorrelated noise produced by dead time, but not on the source-related  signal which is correlated between the two detectors.
This allows to disentangle the effects of dead time from those of the source variability.

@@

\section{Data simulation and analysis} \label{sec:data}
All simulated and real data sets in this paper were produced and/or analyzed with a combination of the two Python libraries \texttt{stingray}%
\footnote{The library is under heavy development.
    For this work we used the version identified by the hash \texttt{XXXXXX}}
\citep{huppenkothen2016} and \texttt{HENDRICS} v.\texttt{3.0b1} \citep[formerly known as MaLTPyNT;][]{2015ascl.soft02021B}, both based on Astropy \citep{astropy2013}.

\subsection{Simulated datasets}
We used the same procedure and algorithms described by \citet{Bachetti+15}, Section 4, that we briefly summarize here together with references to the methods and classes of the libraries above where these steps are implemented.
The classes in \texttt{stingray.simulate} were used to simulate light curves with a given noise profile. 
These classes use the \citet{timmer1995} method to create a noise-affected light curve starting from a given power spectral shape described by a powerlaw index, an \texttt{astropy} model or a list of arrays containing the frequency and power of the model spectrum.
We first created the light curves and then used an acceptance-rejection method to obtain simulated events from these light curves (\texttt{stingray.Eventlist.simulate\_times()}). 
This step needs to be done carefully: if the initial light curves are done directly with the wanted count rate, they contain significant random noise, and the process for the creation of events amplifies this noise (because it further creates a random variate on the top of the local count rate, which is varying randomly already) producing a non-Poissonian final light curve. 
Therefore, we initially used the \citet{timmer1995} method to get light curves with a very high count rate - a way to reduce the relative importance of the Poisson noise because of the square root involved - then renormalized the light curves to the wanted (lower) count rate and rms and finally used these light curves to simulate event lists.
Finally, the \texttt{hendrics.fake.filter\_for\_deadtime()} function was used to apply a non-paralyzable dead time of 2.5\,ms to the simulated event lists. For more details on the simulated data sets, see also Section \ref{sec:correction} and the available Jupyter notebooks%
\footnote{\href{https://github.com/matteobachetti/deadtime-paper-II}{https://github.com/matteobachetti/deadtime-paper-II}} \citep[for a description of Jupyter notebooks, see][]{kluyver2016jupyter}.

\subsection{Cyg X-1 \nustar dataset}
We tested the method also on real data of several bright sources observed by \nustar.
As an example, we report here one such test done on \nustar data of the black hole binary Cyg~X-1 together with the steps to reproduce our analysis, using the linked Jupyter notebook if needed.
We downloaded the observation directory of \nustar ObsID 30001011009 (UT 2014-10-04) from the HEASARC using the custom \texttt{heasarc\_pipelines} package (Bachetti et al., ASCL identifier requested). 
Equivalently, this download can be done using the online tools provided by HEASARC (e.g. \texttt{xamin}).
Starting from the standard cleaned science event files distributed by HEASARC, we applied a barycenter correction using the FTOOL \texttt{barycorr} shipped with HEASOFT 6.21 with the clock correction file n. 71 from the \nustar CALDB.
We selected photons in a region of 50\arcsec around the nominal position of Cyg X-1 using the FTOOL \texttt{fselect}.%
\footnote{To use the method on different sources, it is sufficient to download a different ObsID and change the region file containing the position of the source}.
We used the scripts contained in \texttt{HENDRICS} to load the event lists for the two detectors FPMA and FPMB, calibrate them to translate the PI channels into energy values, and produced light curves with photons from 3 to 79 keV and binned with uniform time bins of 1/4096$\sim2.44\cdot10^{-4}$\,sec%
\footnote{This procedure uses the scripts \texttt{HENreadfile}, \texttt{HENcalibrate} and \texttt{HENlcurve}, as described at\\ \href{http://hendrics.readthedocs.io/en/master/tutorials/quicklook.html}{http://hendrics.readthedocs.io/en/master/tutorials/quicklook.html}}.

@@

\caption{(a) and (b): Real Fourier amplitudes obtained by single light curves (top panels) and difference between two realizations of the same source light curve (bottom) in two cases: (a) Strong $1/f$ red noise and no dead time, calculated over many 500\,s segments of the light curve, and (b) no red noise and strong dead time, calculated over many 5\,s segments of the light curve. 
The choice of different segment length reflects the range of frequencies we want to highlight in the two cases.
The red curve gives the frequency-dependent spread of the distributions, measured by the standard deviation of the curves in each frequency bin. 
As expected, in the first case, the Fourier amplitude follows a power law curve, while the standard deviation of the difference is remarkably stable at all frequencies, as expected by the fact that the Poisson white noise is independent of frequency. 
In the second case, instead, dead time is frequency dependent and white noise is also affected, so that the difference of Fourier amplitudes is modulated as well.
(c) Scatter distribution of the absolute values of dead time-affected Fourier amplitudes versus the difference of Fourier amplitudes for case (b): their relation is clearly linear, with a factor $1/\sqrt{2}$. 
}
\label{fig:fourierdiff}
\end{figure*}

@@

%\begin{figure*}
%\gridline{\fig{imgs/pds_pdf}{0.5\textwidth}{(b)}
%          \fig{imgs/cosp_pdf}{0.5\textwidth}{(c)}}
%
%\caption{(b) Distribution of powers in the periodogram, before the FAD correction and after, shown as a histogram.
%After correction, the powers follow remarkably well the expected $\chi^2_2$ distribution.
%(c) Same, for the cospectrum.
%The correct Laplace distribution is followed after FAD correction}
%\label{fig:dist}
%\end{figure*}
\begin{figure*}
\gridline{\fig{imgs/before_and_after.eps}{\textwidth}{(a)}}
\gridline{\fig{imgs/pds_pdf.pdf}{\textwidth}{(b)}}
\caption{(Top) Periodogram and cospectrum, before and after FAD correction, for a pure white noise light curve. 
The deadtime-driven distortion of the white noise level in the periodogram, and the frequency-dependent modulation of the rms in both spectra, disappear after applying the FAD correction. 
Spectra were calculated over 2-sec intervals and averaged, to decrease the scatter and highlight the distortion of powers.
(Bottom) PDF of non-averaged powers in the cospectrum (red) and the periodogram (black), before the FAD correction and after, shown as a histogram. 
After correction, the powers follow remarkably well the expected Laplace and $\chi^2_2$ distributions respectively, as highlighted by the overplotted probability density functions (PDF).}
\label{fig:dist}
\end{figure*}
Let us simulate two constant light curves with an incident mean count rate of 400 counts/sec and a dead time of 2.5 ms, as we would expect from two identical detectors observing the same stable X-ray source. 
This case is illustrated in Figure ~\ref{fig:fourierdiff} (right panel).
The Fourier amplitudes $A_{xj}$ and $A_{yj}$ of the light curves from the two detectors are heavily distorted by dead time, with the characteristic damped oscillator-like shape \citep{Vikhlinin+94,Zhang+95}. 
As laid out in Section \ref{sec:fourierdiff}, the difference of Fourier amplitudes from two independent but identical detectors shows no source variability, but \textit{still shows the same distortion} due to dead time.
This gives a clear way to disentangle between source- and dead time-driven variability.
By using the difference between the Fourier amplitudes in two detectors, we can in principle renormalize the power spectrum so that only the source variability alters its otherwise flat shape.

As shown in Figure~\ref{fig:dist} (left panel), the single-channel Fourier amplitudes are proportional to the difference of the Fourier amplitudes in different realizations, with a constant factor $1/\sqrt{2}$.
Therefore, we expect that the periodogram will be proportional to the square of the Fourier amplitude difference, divided by 2.
Let us try to \textit{divide the power spectrum by a smoothed version of the squared Fourier differences}, and multiply by 2.
For smoothing, we used a Gaussian running window with a window width of 50 bins.
Given that the initial binning had 50 bins/Hz, this interpolation allows an aggressive smoothing over bins whose y value does not change significantly.
In general, we recommend smoothing over as many bins as allowed by the shape of the periodogram.
In this paper, when not specified we average over the number of bins contained in 1Hz of the spectrum.
We call this procedure the \textbf{Fourier Amplitude Difference} (hereafter FAD) \textbf{correction}.

The correction is depicted in Figure~\ref{fig:dist}, right. 
Starting from a heavily distorted distribution of the powers, applying the FAD correction reinstates a remarkably correct distribution of powers, following the expected $\chi^2_2$ distribution \citep{Lewin+88} very closely.
Analogously, the corrected cospectrum will follow the expected Laplace distribution.
While the original dead time-affected cospectrum had a frequency-dependent modification to the rms level, the FAD-corrected cospectrum gets back to a frequency-independent shape, like in the dead time-free case.

@@

\section{Testing the FAD correction on simulated data} \label{sec:correction}

%\begin{figure*}
%    \gridline{\fig{imgs/rednoise_rebin}{0.33\textwidth}{(a)}
%        \fig{imgs/rednoise_rebin_zoom}{0.33\textwidth}{(b)}
%        \fig{imgs/simu_rebin_log}{0.33\textwidth}{(c)}}
%    \gridline{\fig{imgs/cygx1_soft_nustar_2_rebin}{0.33\textwidth}{(d)}
%        \fig{imgs/cygx1_soft_nustar_2_rebin_zoom}{0.33\textwidth}{(e)}
%        \fig{imgs/cygx1_soft_nustar_2_rebin_log}{0.33\textwidth}{(f)}}  
%    \caption{
%        Top: simulated dataset with two strong and broad Lorentzian components, incident count rate $\sim$810 ct/s, and total rms $\sim$15\% (``detected'' after dead time: $\sim$270 and $\sim$4\% resp.).
%        Bottom: \nustar observation of the black hole candidate Cyg X-1. 
%        (a) and (d): comparison of the distortion of the dead time-affected periodogram with the FAD-corrected periodogram, the dead time-affected and the FAD-corrected cospectrum. The spectra have been shifted vertically for clarity.
%        The shape of the uncorrected periodogram is clearly distorted by dead time. 
%        (b) and (e): the FAD correction successfully flattens the periodogram, but there remain a few intervals where the baseline is imperfect, indicated by the arrows.
%        This is due to the flux mismatch between the two detectors (see \sref{sec:realdata}).
%        The cospectral powers are always distributed around zero, but dead time changes the width of the resulting distribution, and the FAD correction attenuates the width of the distribution back to the width expected in the dead time-free case (see e.g.\ (b) for a striking case).
%        Moreover, the source-dominated part of the spectrum also receives a boost, which corrects precisely the measured rms of the source from the ``hushing'' effect of dead time.
%        (c) and (f): FAD-corrected periodogram and cospectra, plotted in $\mathrm{(rms/mean)}^2$ normalization and fitted with two Lorentzian curves (c) and with an exponential cutoff power law (f).
%        The additional gain coming from the FAD normalization is evident by the comparison with the uncorrected cospectrum.}
%    \label{fig:pds}
%\end{figure*}
\begin{figure*}
    \plotone{imgs/four_qpos.eps}
    \caption{Periodograms and cospectra from a simulation with four Lorentzian features (at 50, 200, 300 and 400 Hz) with 40-Hz full with at half maximum (FWHM). 
    We plotted and fitted periodograms and cospectra before and after applying the dead time filter
    The total rms before dead time was 40\% and the incident photon flux 400 ct/s. 
    After applying a dead time of 2.5 ms like in \nustar, the ``detected'' photon flux decreased to $\sim$200 ct/s as expected. 
    As can be seen from the best-fit curves, there is no significant difference between FAD-normalized and deadtime-free periodograms and cospectra.
    For comparison, we also plot the cospectrum without FAD (pink), showing very different amplitudes for the four Lorentzians, due to dead time.
    }
    \label{fig:pds}
\end{figure*}

We are now ready to verify the last step: is the FAD-corrected power spectrum equivalent (albeit with some loss of sensitivity due to the lower number of photons) to the dead time-free power spectrum?
To test this, we produced a number of different synthetic datasets as explained in Section~\ref{sec:wndeadtime}, containing different combinations of QPOs and broadband noise components. 
We first calculated the periodogram and cospectrum of the dead time-free data, averaged over 128-s intervals.
Then, we applied a dead time filter to the event list and applied the FAD correction, as follows:
\begin{enumerate}
\item split the two light curves in segments of 128 to 512 seconds
\item for each pair of light curve segments: 
	\begin{itemize}
	\item calculate the Fourier transform of each channel separately, and then of the summed channels (hereafter the \textit{total-intensity} light curve);
	\item multiply the Fourier amplitudes by $\sqrt{2/N_{ph}}$ in order to obtain Leahy-normalized periodograms;
	\item subtract the Fourier amplitudes of the two channels between them, \textit{take the absolute value}, and obtain this way the Fourier Amplitude Differences (FAD);
	\item \textit{smooth} the FAD using a Gaussian-window interpolation with a width of 1-2 seconds;
	\item use the separated single-channel and total-intensity Fourier amplitudes to calculate the periodograms;
	\item use the Fourier amplitudes from channels A and B to calculate the cospectrum;
	\item divide all periodograms and the cospectrum by the smoothed and squared FAD, and multiply by 2.
	\end{itemize}
\end{enumerate}

All spectra were then expressed in fractional rms \citep{BelloniHasinger90,Miyamoto+91} normalization, where the integral of the fitted spectral components over the frequency returns directly its fractional rms.
In the rms normalization, the values of each point of the periodogram, after white noise subtraction, should be consistent between the dead time-free and the FAD-corrected periodograms.
We first checked visually that the spectra (white-noise subtracted in the case of periodograms) after FAD correction were all consistent with the dead time-free periodogram. 
Then, we fitted all spectra with the model which produced the simulated data and checked if the values were consistent with the input model parameters and the fit on the dead time-free periodogram.
Finally, we verified if \textit{at low frequencies} the total rms of the FAD-corrected spectra was consistent with the total rms of the dead time-affected cospectrum \textit{times the ratio between incident and detected photons}, using the count rates before and after applying the dead time filter.
To calculate this rms, we fitted the spectra with two Lorentzian components, and integrated the model over the full frequency range.
For periodograms, the model included also a constant offset to account for the white noise level.
An example of this analysis is shown in Figure~\ref{fig:pds}.

\section{Caveats}\label{sec:caveat}
The simulations show that the shape of the periodogram is precisely corrected by the FAD procedure \textit{if the input light curves have the same count rate} and \textit{for values of the input count rate and rms that are not too extreme}. 
The first caveat turns out to be an issue only for the single-channel periodogram. 
At high count rates, single-channel periodograms are corrected very well only if the two channels have very similar count rates, with the requirement being unsurprisingly stricter and stricter as the count rate increases.
However, we find that the total-intensity periodogram and the cospectrum remains well corrected by the FAD even if the count rate in the two channels differs by 30\%, for any reasonable count rate.
Therefore, we recommend to use the FAD very carefully with single-channel periodograms, which should not be an issue given that the total-intensity periodogram is more sensitive and more convenient to use anyway. 
A comparison with the cospectrum, that is not affected by white noise distortions, is always recommended.

The second caveat needs some attention. 
We find that the FAD correction consistently overestimates the integrated rms when the count rate and rms are \textit{both} very high, \textit{independently from the frequency}.
This latter feature is less intuitive, because we would expect that this overestimation should somehow be more relevant where the correction is stronger, while this is not the case.
But for this very reason, it is somewhat easier to account for this through simulations, without having to make further risky frequency-dependent corrections to the rms.
At standard count rates and fractional rms for bright black hole binaries, $\sim$200 ct/s and 30\% rms, the difference is below 5\%. 
However, it increases \textit{linearly} with the count rate and \textit{quadratically} with rms.
A practical way to estimate this effect during analysis is to apply the FAD, obtain a best fit model, calculate the rms, then use the codes we shared with the Jupyter notebooks to simulate a number of realizations of the code, evaluating what is the amount of overestimation involved.

>>>>

\newcommand{\given}{\ensuremath{\,|\,}\xspace}
\newcommand{\dd}{\ensuremath{\mathrm{d}}\xspace}
\newcommand{\counts}{\ensuremath{y}\xspace}
\newcommand{\rms}{\ensuremath{\mathrm{r.m.s.}}\xspace}
\newcommand{\pars}{\ensuremath{\theta}\xspace}
\newcommand{\mean}{\ensuremath{\lambda}\xspace}
\newcommand{\likelihood}{\ensuremath{{\mathcal L}}\xspace}
\newcommand{\Poisson}{\ensuremath{{\mathcal P}}\xspace}
\newcommand{\Uniform}{\ensuremath{{\mathcal U}}\xspace}

\newcommand{\Normal}{\ensuremath{{\mathcal N}}\xspace}
\newcommand{\bg}{\ensuremath{\mathrm{bg}}\xspace}
\newcommand{\word}{\ensuremath{\phi}\xspace}

@@

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
Dead time affects many of the instruments used in X-ray astronomy, by producing a strong distortion in power density spectra. This can make it difficult to model the aperiodic variability of the source or look for quasi-periodic oscillations.
Whereas in some instruments a simple a-priori correction for dead time-affected power spectra is possible, 
this is not the case for others such as \nustar, where the dead time is non-constant and long ($\sim$2.5\,ms).
\citet{Bachetti+15} suggested the cospectrum obtained from light curves of independent detectors within the same instrument as a possible way out, but this solution has always only been a partial one: the measured \rms was still affected by dead time, because the width of the power distribution of the cospectrum was modulated by dead time in a frequency-dependent way.

@@

When using data from missions carrying two or more \textit{identical and independent} detectors like \nustar, \citet{Bachetti+15} proposed an approach to mitigate instrumental effects like dead time exploiting this redundancy: 
where in standard analysis, light curves of multiple detectors are summed before Fourier transforming the summed light curve, 
it is possible to instead Fourier-transform the signal of two independent detectors and combine the Fourier amplitudes in a \textit{cospectrum} -- the real part of the cross spectrum -- instead of the periodogram. 
Since dead time is uncorrelated between the two detectors, the resulting powers have a mean white noise level fixed to 0, which resolves the first and most problematic issue created by dead time (see details in \citealt{Bachetti+15}); however, the resulting powers no longer follow the statistical distribution expected for power spectra, and their probability distribution is frequency-dependent.
Whereas a noise cospectrum in the absence of dead time would follow a Laplace distribution (Huppenkothen and Bachetti, sub.),
dead time affects the width of the probability distribution for cospectral powers and modulates the measured \rms proportionally to the distortion acted on power spectra.
In this Letter, we show a method to precisely recover the shape of the power density spectrum by looking at the difference of the Fourier amplitudes of the light curves of two independent detectors.
This difference, in fact, contains information on the uncorrelated noise produced by dead time, but not on the source-related  signal which is correlated between the two detectors.
This allows to disentangle the effects of dead time from those of the source variability.

@@

\section{Data simulation} \label{sec:data}
All simulated and real data sets in this paper were produced and/or analyzed with a combination of the two Python libraries \texttt{stingray}%
\footnote{The library is under heavy development.
    For this work we used the version identified by the hash \texttt{XXXXXX}}
\citep{huppenkothen2016} and \texttt{HENDRICS} v.\texttt{3.0b1} \citep[formerly known as MaLTPyNT;][]{2015ascl.soft02021B}, both based on Astropy \citep{astropy2013}.

We used the same procedure and algorithms described by \citet{Bachetti+15}, Section 4, that we briefly summarize here together with references to the methods and classes of the libraries above where these steps are implemented.
The classes in \texttt{stingray.simulate} were used to simulate light curves with a given noise profile. 
These classes use the \citet{timmer1995} method to create a noise-affected light curve starting from a given power spectral shape.
We first created the noisy light curves and then used an acceptance-rejection method to obtain simulated events from these light curves (\texttt{stingray.Eventlist.simulate\_times()}). 
This step needs to be done carefully: if the initial light curves are done directly with the wanted count rate, they contain significant random noise, and the process for the creation of events amplifies this noise (because it further creates a random variate on the top of the local count rate, which is varying randomly already) producing a non-Poissonian final light curve. 
Therefore, we initially used the \citet{timmer1995} method to get light curves with a very high count rate - so that the Poisson noise is relatively small - then renormalized the light curves to the wanted (lower) count rate and \rms and finally used these light curves to simulate event lists.
Finally, the \texttt{hendrics.fake.filter\_for\_deadtime()} function was used to apply a non-paralyzable dead time of 2.5\,ms to the simulated event lists. For more details on the simulated data sets, see also Section \ref{sec:correction} and the available Jupyter notebooks%
\footnote{\href{https://github.com/matteobachetti/deadtime-paper-II}{https://github.com/matteobachetti/deadtime-paper-II}} \citep[for a description of Jupyter notebooks, see][]{kluyver2016jupyter}.

%\subsection{Cyg X-1 \nustar dataset}
%We tested the method also on real data of several bright sources observed by \nustar.
%As an example, we report here one such test done on \nustar data of the black hole binary Cyg~X-1 together with the steps to reproduce our analysis, using the linked Jupyter notebook if needed.
%We downloaded the observation directory of \nustar ObsID 30001011009 (UT 2014-10-04) from the HEASARC using the custom \texttt{heasarc\_pipelines} package (Bachetti et al., ASCL identifier requested). 
%Equivalently, this download can be done using the online tools provided by HEASARC (e.g. \texttt{xamin}).
%Starting from the standard cleaned science event files distributed by HEASARC, we applied a barycenter correction using the FTOOL \texttt{barycorr} shipped with HEASOFT 6.21 with the clock correction file n. 71 from the \nustar CALDB.
%We selected photons in a region of 50\arcsec around the nominal position of Cyg X-1 using the FTOOL \texttt{fselect}.%
%\footnote{To use the method on different sources, it is sufficient to download a different ObsID and change the region file containing the position of the source}.
%We used the scripts contained in \texttt{HENDRICS} to load the event lists for the two detectors FPMA and FPMB, calibrate them to translate the PI channels into energy values, and produced light curves with photons from 3 to 79 keV and binned with uniform time bins of 1/4096$\sim2.44\cdot10^{-4}$\,sec%
%\footnote{This procedure uses the scripts \texttt{HENreadfile}, \texttt{HENcalibrate} and \texttt{HENlcurve}, as described at\\ \href{http://hendrics.readthedocs.io/en/master/tutorials/quicklook.html}{http://hendrics.readthedocs.io/en/master/tutorials/quicklook.html}}.

@@

\caption{(a) and (b): Real Fourier amplitudes obtained by single light curves (top panels) and difference between two realizations of the same source light curve (bottom) in two cases: (a) Strong $1/f$ red noise and no dead time, calculated over many 500\,s segments of the light curve, and (b) no red noise and strong dead time, calculated over many 5\,s segments of the light curve. 
The choice of different segment length reflects the range of frequencies we want to highlight in the two cases.
The red curve gives the frequency-dependent spread of the distributions, measured by the mean of the absolute values of the curves in each frequency bin. 
As expected, in the first case, the Fourier amplitude follows a power law curve, while the standard deviation of the difference is remarkably stable at all frequencies, as expected by the fact that the Poisson white noise is independent of frequency. 
In the second case, instead, dead time is frequency dependent and white noise is also affected, so that the difference of Fourier amplitudes is modulated as well.
(c) Scatter distribution of the absolute values of dead time-affected Fourier amplitudes versus the difference of Fourier amplitudes for case (b): their relation is clearly linear, with a factor $1/\sqrt{2}$. 
}
\label{fig:fourierdiff}
\end{figure*}

@@

\begin{figure*}
\plotone{imgs/before_and_after.eps}
%\plotone{imgs/pds_pdf.pdf}
\caption{Periodogram and cospectrum, before and after FAD correction, for a pure white noise light curve. 
The deadtime-driven distortion of the white noise level in the periodogram, and the frequency-dependent modulation of the \rms in both spectra, disappear after applying the FAD correction. 
Spectra were calculated over 2-sec intervals and averaged, to decrease the scatter and highlight the distortion of powers.}
\label{fig:comparison}
\end{figure*}

\begin{figure*}
\plotone{imgs/pds_pdf.pdf}
\caption{PDF of non-averaged powers in the cospectrum (red) and the periodogram (black), before the FAD correction and after, shown as a histogram. 
After correction, the powers follow remarkably well the expected Laplace and $\chi^2_2$ distributions respectively, as highlighted by the overplotted probability density functions (PDF).}
\label{fig:dist}
\end{figure*}

Let us simulate two constant light curves with an incident mean count rate of 400 counts/sec [CHECK] and a dead time of 2.5 ms, as we would expect from two identical detectors observing the same stable X-ray source. 
This case is illustrated in Figure ~\ref{fig:fourierdiff} (middle panel).
The Fourier amplitudes $A_{xj}$ and $A_{yj}$ of the light curves from the two detectors are heavily distorted by dead time, with the characteristic damped oscillator-like shape \citep{Vikhlinin+94,Zhang+95}. 
As laid out in Section \ref{sec:fourierdiff}, the difference of Fourier amplitudes from two independent but identical detectors shows no source variability, but \textit{still shows the same distortion} due to dead time.
This gives a clear way to disentangle between source- and dead time-driven variability.
By using the difference between the Fourier amplitudes in two detectors, we can in principle renormalize the power spectrum so that only the source variability alters its otherwise flat shape.

As shown in Figure~\ref{fig:fourierdiff} (left panel), the single-channel Fourier amplitudes are proportional to the difference of the Fourier amplitudes in different realizations, with a constant factor $1/\sqrt{2}$.
Therefore, we expect that the periodogram will be proportional to the square of the Fourier amplitude difference, divided by 2.
Let us try to \textit{divide the power spectrum by a smoothed version of the squared Fourier differences}, and multiply by 2.
For smoothing, we used a Gaussian running window with a window width of 50 bins.
Given that the initial binning had 50 bins/Hz, this interpolation allows an aggressive smoothing over bins whose y value does not change significantly.
We call this procedure the \textbf{Fourier Amplitude Difference} (hereafter FAD) \textbf{correction}.

The correction is depicted in Figure~\ref{fig:comparison}. 
Starting from a heavily distorted distribution of the powers, applying the FAD correction ``flattens'' remarkably well the white noise level of the periodogram and the distribution of the scatter of the white noise periodogram and cospectrum. 
Also, it reinstates a remarkably correct distribution of powers, following the expected $\chi^2_2$ distribution \citep{Lewin+88} very closely. 
Analogously, the corrected cospectrum will follow the expected Laplace distribution \citep{Huppenkothen & Bachetti, sub.}.
While the original dead time-affected cospectrum had a frequency-dependent modification to the \rms level, the FAD-corrected cospectrum gets back to a frequency-independent shape, like in the dead time-free case.

@@

\section{The FAD correction algorithm} \label{sec:fad}
In practice, the FAD correction algorithm in a generic case would work as follows:
\begin{enumerate}
\item split the two light curves in segments as one would do to calculate standard averaged periodograms;
\item for each pair of light curve segments: 
	\begin{itemize}
	\item calculate the Fourier transform of each channel separately, and then of the summed channels (hereafter total-intensity);
	\item save the unnormalized Fourier amplitudes;
	\item multiply these Fourier amplitudes by $\sqrt{2/N_{ph}}$ (that would give Leahy-normalized periodograms if squared);
	\item subtract the Leahy-normalized Fourier amplitudes of the two channels between them, \textit{take the absolute value}, and obtain this way the Fourier Amplitude Differences (FAD);
	\item \textit{smooth} the FAD using a Gaussian-window interpolation with a large number of bins, in our case all the bins contained in 1-2 Hz;
	\item use the separated single-channel and total-intensity \textit{unnormalized} Fourier amplitudes to calculate the periodograms;
	\item use the unnormalized Fourier amplitudes from channels A and B to calculate the cospectrum;
	\item divide all periodograms and the cospectrum by the smoothed and squared FAD, and multiply by 2.
	\end{itemize}
\item normalize the periodograms to the wanted normalization (Leahy or fractional \rms).
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%% SECTION 6 %%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Testing the FAD correction on simulated data} \label{sec:correction}

\begin{figure*}
    \plotone{imgs/four_qpos.eps}
    \caption{Periodograms and cospectra from a simulation with four Lorentzian features (at 50, 200, 300 and 400 Hz) with 40-Hz full with at half maximum (FWHM). 
    We plotted and fitted periodograms and cospectra before and after applying the dead time filter
    The total \rms before dead time was 40\% and the incident photon flux 400 ct/s. 
    After applying a dead time of 2.5 ms like in \nustar, the ``detected'' photon flux decreased to $\sim$200 ct/s as expected. 
    As can be seen from the best-fit curves, there is no significant difference between FAD-normalized and deadtime-free periodograms and cospectra.
    For comparison, we also plot the cospectrum without FAD (pink), showing very different amplitudes for the four Lorentzians, due to dead time.
    }
    \label{fig:pds}
\end{figure*}

We are now ready to verify the last step: is the FAD-corrected power spectrum equivalent (albeit with some loss of sensitivity due to the lower number of photons) to the dead time-free power spectrum?
To test this, we produced a number of different synthetic datasets as explained in Section~\ref{sec:wndeadtime}, containing different combinations of QPOs and broadband noise components. 
We first calculated the periodogram and cospectrum of the dead time-free data, averaged over 128-s intervals.
Then, we applied a dead time filter to the event list and applied the FAD correction, as described in Section~\ref{sec:fad}

All spectra were then expressed in fractional \rms \citep{BelloniHasinger90,Miyamoto+91} normalization, where the integral of the fitted spectral components over the frequency returns directly its fractional \rms.
In the \rms normalization, the values of each point of the periodogram, after white noise subtraction, should be consistent between the dead time-free and the FAD-corrected periodograms.
We fitted all spectra with the model which produced the simulated data and checked if the \rms values were consistent with the fit on the dead time-free periodogram.
To calculate this \rms, we fitted the spectra with two Lorentzian components, and integrated the model over the full frequency range.
For periodograms, the model included also a constant offset to account for the white noise level.
An example of this analysis is shown in Figure~\ref{fig:pds}.

\section{Caveats}\label{sec:caveat}
\begin{figure*}
    \plotone{imgs/errors.png}
    \caption{Relative overestimation of FAD with respect to \rms, versus \rms, as calculated from the cospectrum.  
    We encoded the frequency of the feature in the color, and the incident rate in the size of the scatter points.
    From this visualization we see two regimes: below $\sim$40\% fractional \rms, the errors are dominated by statistical errors. 
    These errors will simply decrease when we average more data, as we expect from statistical errors.
    Over $\sim$40\% fractional \rms, the errors are significantly skewed towards an overestimation of the \rms, and this is 
    clearly truer when the incident rate \textit{and} the \rms are high.}
    \label{fig:errors}
\end{figure*}
The simulations show that the shape of the periodogram is precisely corrected by the FAD procedure \textit{if the input light curves have the same count rate} and \textit{for values of the input count rate and \rms that are not too extreme}. 
The first caveat turns out to be an issue only for the single-channel periodogram. 
At high count rates, single-channel periodograms are corrected very well only if the two channels have very similar count rates, with the requirement being unsurprisingly stricter and stricter as the count rate increases.
However, we find that the total-intensity periodogram and the cospectrum remains well corrected by the FAD even if the count rate in the two channels differs by 30\%, for any reasonable count rate.
Therefore, we recommend to use the FAD very carefully with single-channel periodograms, which should not be an issue given that the total-intensity periodogram is more sensitive and more convenient to use anyway. 
A comparison with the cospectrum, that is not affected by white noise distortions, is always recommended.

The second caveat needs some attention. 
We find that the FAD correction consistently overestimates the integrated \rms when the count rate and \rms are \textit{both} very high, \textit{in particular at low frequency} (See Figure~\ref{fig:errors}).
At standard count rates and fractional \rms for bright black hole binaries, $\sim$200 ct/s and 30\% \rms for the \nustar case, the relative error is below 5\% (meaning that if the \rms is 30\%, the measured \rms is between 30 and 32\%!) and it is symmetrically distributed around 0, as expected from statistical error. 
At higher incident rates and \rms, the error distribution skews towards positive relative errors, implying an overestimation of the \rms.
This should not be a problem in most cases, where the \rms is used as a rough indicator for spectral state.
If very precise measurements of \rms are needed (for example, to calculate \rms/energy spectra), it is somewhat safer to account for this through simulations, without having to make further risky frequency-dependent corrections to the \rms.
As a rough estimate, the relative error on the fractional \rms increases \textit{linearly} with the count rate and \textit{quadratically} with real \rms.
A practical way to estimate this effect during analysis is to apply the FAD, obtain a best fit model, calculate the \rms, then use the codes we shared with the Jupyter notebooks to simulate a number of realizations of the light curve, evaluating what is the amount of overestimation involved.
